---
title: "D208 - Predictive Modeling"
output: html_notebook
---
___

#### **Performance Assessment - Task 1: Multiple Regression for Predictive Modeling**
*Medical Readmission Data Set (Clean)*

___ 


</br>


## Part I

### A1: Research Question

### A2: Objectives and Goals


</br>


## Part II

### B1: Summary of Assumptions

### B2: Tool Benefits

### B3: Appropriate Technique


</br>


## Part III

### C1: Data Goals

### C2: Summary Statistics

### C3: Steps to Prepare the Data

### C4: Visualizations

### C5: Prepared Data Set


</br>


## Part IV

### D1: Initial Model

### D2: Justification of Model Reduction

### D3: Reduced Multiple Regression Model


</br>


## Part V

### E1: Model Comparison

### E2: Output and Calculations

### E3: Code
```{r}
# Load in the tidyverse
# Suppress the overwritten functions messages
# suppressPackageStartupMessages(library(tidyverse))
# Because the above caused an error with the lsp, use for final draft
library(tidyverse)
```

```{r}
# Read dataset in
df <- read.csv("./data/medical_clean.csv",
               header = TRUE)
```

```{r}
# Start with a quick glimpse of the data for orientation and future reference
glimpse(df)
```

```{r}
# Partition dataset to include only features to be initially included in model
df_mdl <- df %>%
  select(Initial_days,
         Age,
         Gender:Services)
```

```{r}
df_mdl %>%
  glimpse()
```

```{r}
# Partition continuous vars from df_mdl
df_mdl_doubles <- df_mdl %>%
  select_if(is.double)
```

```{r}
# use Hmisc package to view hist matrix of dbl vars
library(Hmisc)
df_mdl_doubles %>%
  hist.data.frame()
```

```{r}
init_mdl <- lm(Initial_days ~
               Age +
               Gender +
               ReAdmis +
               VitD_levels +
               Doc_visits +
               Full_meals_eaten +
               vitD_supp +
               Soft_drink +
               Initial_admin +
               HighBlood +
               Stroke +
               Complication_risk +
               Overweight +
               Arthritis +
               Diabetes +
               Hyperlipidemia +
               BackPain +
               Anxiety +
               Allergic_rhinitis +
               Reflux_esophagitis +
               Asthma +
               Services,
               data = df_mdl)

summary(init_mdl)
```

```{r}
df_mdl_emergadmin <- df_mdl %>%
    filter(Initial_admin == "Emergency Admission")
```

```{r}
reduced_mdl <- lm(Initial_days ~
                  VitD_levels +
                  0 +
                  ReAdmis,
                  data = df_mdl_emergadmin)

summary(reduced_mdl)
```


</br>


## Part VI

### F1: Results

### F2: Recommendations


</br>


### I: Sources

