---
title: "D208 - Predictive Modeling"
output: html_notebook
---
___

#### **Performance Assessment - Task 1: Multiple Regression for Predictive Modeling**
*Medical Readmission Data Set (Clean)*

___ 


</br>


## Part I

### A1: Research Question
What variables from the medical dataset are related to and influence a patient's initial number of days spent after hospitalization? In terms of hypothesis testing, our null hypothesis (H0) is, "No combination of variables included in the medical dataset are related to initial length of stay in any statistically significant way. Additionally, our alternate hypothesis (H1) is, "Some combination of variables included in the medical dataset are related to initial length of stay in a statistically significant way". 

### A2: Objectives and Goals
The main goal of the following analysis is to determine what variables, if any, contribute positively to a patient's initial number of days hospitalized. This will be assessed using the R programming language and the technique of multiple regression to identify a meaningful model.

</br>


## Part II

### B1: Summary of Assumptions
The following are basic assumptions of multiple linear regression:

- The assumption of a linear relationship between the explanatory variables and the target variable
- The assumption that residual values are normally distributed
- The assumption of non-multicollinearity of explanatory variables
- The assumption that residual values are homoscedastic

### B2: Tool Benefits
The programming language of choice for this analysis will be the R programming language. Previously, Python has been the language most utilized, however R happens to handle the tasks of regression exceptionally well and thus became the self-evident choice for multiple regression analysis. In particular, the built-in functions of the base R language used to fit models are incredibly simple to navigate as well as demonstrate. Additionally, the reduction process using the `ols_step_backward_p()` function in particular made the process even easier. It is for good reason that R has a reputation for handling regression models.

### B3: Appropriate Technique
Multiple regression is an appropriate technique to use to accomplish our goal of finding which variable contribute to a longer initial length of stay for several reasons. Firstly, the dataset contains 50 total variables of varying significance and utility. While some variable do not require much thought prior to elimination, some are not quite as straightforward. Therefore, running them through a multiple regression model prior to discarding them is a worthwhile venture. Additionally, multiple regression will allow us to see the significance of each variable's discrete contribution to the target variable as well as the interaction between explanatory variables themselves. Overall, regarding our objective, multiple regression is more than up to the task.

</br>


## Part III

### C1: Data Goals

### C2: Summary Statistics

### C3: Steps to Prepare the Data

### C4: Visualizations

### C5: Prepared Data Set


</br>


## Part IV

### D1: Initial Model

### D2: Justification of Model Reduction

### D3: Reduced Multiple Regression Model


</br>


## Part V

### E1: Model Comparison

### E2: Output and Calculations

### E3: Code
```{r}
# Load in the tidyverse
# Suppress the overwritten functions messages
# suppressPackageStartupMessages(library(tidyverse))
# Because the above caused an error with the lsp, use for final draft
library(tidyverse)
```

```{r}
# Read dataset in
df <- read.csv("./data/medical_clean.csv",
               header = TRUE)
```

```{r}
# Start with a quick glimpse of the data for orientation and future reference
glimpse(df)
```

```{r}
# Partition dataset to include only features to be initially included in model
df_mdl <- df %>%
  select(Initial_days,
         Age,
         Gender:Services)
```

```{r}
df_mdl %>%
  glimpse()
```

```{r}
# Partition continuous vars from df_mdl
df_mdl_doubles <- df_mdl %>%
  select_if(is.double)
```

```{r}
# use Hmisc package to view hist matrix of dbl vars
library(Hmisc)
df_mdl_doubles %>%
  hist.data.frame()
```

```{r}
init_mdl <- lm(Initial_days ~
               Age +
               Gender +
               ReAdmis +
               VitD_levels +
               Doc_visits +
               Full_meals_eaten +
               vitD_supp +
               Soft_drink +
               Initial_admin +
               HighBlood +
               Stroke +
               Complication_risk +
               Overweight +
               Arthritis +
               Diabetes +
               Hyperlipidemia +
               BackPain +
               Anxiety +
               Allergic_rhinitis +
               Reflux_esophagitis +
               Asthma +
               Services,
               data = df_mdl)

summary(init_mdl)
```

```{r}
df_mdl_emergadmin <- df_mdl %>%
    filter(Initial_admin == "Emergency Admission")
```

```{r}
reduced_mdl <- lm(Initial_days ~
                  VitD_levels +
                  0 +
                  ReAdmis,
                  data = df_mdl_emergadmin)

summary(reduced_mdl)
```


</br>


## Part VI

### F1: Results

### F2: Recommendations


</br>


### I: Sources

